Foregrounds in Fourier Space for Redshifted 21~cm Observatories

### 2015-02-04 ms.tex
Add a new section "Wide-field measurements"
from cfc6578:
============

In ms.tex,

(*) Add a new section "Wide-field measurements"
(*) Need to add a figure showing wide-field effects

In \S\ref{sec:delay-spectrum} we {\bf provide a brief overview of the delay spectrum technique} \citep{par12a,par12b}. In \S\ref{sec:instrument}, we describe the MWA setup, summarize the observing parameters, and present the resulting data. Simulations using these observing parameters are described in \S\ref{sec:modeling} and analyzed by sky location and type of emission in \S\ref{sec:delay-spectrum-analysis}. In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique. {\bf The role of antenna shapes in determining foreground contamination is explored in \S\ref{sec:antenna-shape}}. We present a summary of our work and findings in \S\ref{sec:summary}.

@@

For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all baselines to lie in the range $-67.5\arcdeg \leq \theta_b < 112.5\arcdeg$. Baselines oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis. {\bf Note that visibilities from different baselines have not been averaged together in this analysis}.

\section{Example Data}\label{sec:instrument}

\subsection{The Murchison Widefield Array (MWA)}

>>>>

In \S\ref{sec:delay-spectrum} we {\bf provide a brief overview of the delay spectrum technique} \citep{par12a,par12b}. In \S\ref{sec:data}, we describe the MWA setup, summarize the observing parameters, and present the resulting data. Simulations using these observing parameters are described in \S\ref{sec:modeling} and analyzed by sky location and type of emission in \S\ref{sec:delay-spectrum-analysis}. In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique. {\bf The role of antenna shapes in determining foreground contamination is explored in \S\ref{sec:antenna-shape}}. We present a summary of our work and findings in \S\ref{sec:summary}.

@@

In our present study, we use an antenna layout identical to that of the MWA \citep{bea12}. The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km. Here we focus on antenna spacings $|\boldsymbol{b}| \le 200$~m (spatial scales relevant to reionization). For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all baselines to lie in the range $-67.5\arcdeg \leq \theta_b < 112.5\arcdeg$. Baselines oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis. {\bf Note that visibilities from different baselines have not been averaged together in this analysis}.

\section{Wide--Field Measurements}\label{sec:wide-field}

With $\hat{\boldsymbol{s}}\equiv (l,m,n)$, equation~\ref{eqn:obsvis} can be written as \citep{tay99,tho01}:
\begin{align}\label{eqn:obsvis-lmn}
  V_{bf} &= \iint\limits_\textrm{sky} \frac{A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)}{\sqrt{1-l^2-m^2}}\,W_f\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif l\,\dif m, 
\end{align}
where, $l$, $m$, and $n$ denote the direction cosines toward east, north, and zenith respectively, with $n\equiv\sqrt{1-l^2-m^2}$, and:
\begin{align}\label{eqn:solid-angle}
  \dif\Omega = \frac{\dif l\,\dif m}{\sqrt{1-l^2-m^2}}.
\end{align}

When the synthesized field is small, where $A(\hat{\boldsymbol{s}},f)$ or $I(\hat{\boldsymbol{s}},f)$ is negligible for $|l| \ll 1$ and $|m| \ll 1$, equation~\ref{eqn:obsvis-lmn} reduces to a simple two--dimensional Fourier transform \citep{tay99,tho01} between the apparent sky brightness and measured visibilities. It is in this context that radio interferometers are understood to be sensitive only to fluctuations and not to a uniform sky brightness distribution.

In a wide--field measurement, neither $A(\hat{\boldsymbol{s}},f)$ nor $I(\hat{\boldsymbol{s}},f)$, in general, is negligible anywhere in the visible hemisphere. The solid angle per pixel on the sky in direction cosine coordinates changes significantly with direction (equation~\ref{eqn:solid-angle}), increasing steeply towards the horizon. Hence, the approximations in the narrow--field scenario do not apply. For example, even if $A(\hat{\boldsymbol{s}},f)$ and $I(\hat{\boldsymbol{s}},f)$ are held constant across the visible hemisphere, the integrand in equation~\ref{eqn:obsvis-lmn} is still dependent on direction. Therefore, in a significant departure from a narrow--field measurement, the wide--field visibility from a uniform brightness distribution on a non--zero antenna spacing is not zero.

Figure ??? shows the wide--field delay power spectrum response of a uniformly illuminated sky with no spectral variation measured by antenna elements with a uniform power pattern across the sky. Notice the features arising due to the direction--dependent solid angle term and the steep rise in power towards the horizon limits. Wide--field effects are prevalent on all antenna spacings, including the longest ones used in this study.

\citet{thy13} have found evidence of this feature in their statistical models. In line with our reasoning, they attribute it to an increase in solid angles subtended by delay bins closer to the horizon limits thereby, enhancing the resulting integrated emission. \citet{pob13} also find a similar ``edge brightening'' feature which they attribute to emission from the Galactic plane near the horizon. From their discussion, it is unclear what fraction of power in that feature arises from such wide--field effects.

We conclude these are generic to all instruments making wide--field measurements. The nature of the specific instrument used for observing will control the amplitude of these effects, which we explore below. % Typically, the foreground sky consists of diffuse emission and point-like sources, and the antenna is more sensitive to power received from primary field of view than from the rest of the sky.

\section{Simulations}\label{sec:sim}

\section{Example Data}\label{sec:data}

\subsection{The Murchison Widefield Array (MWA)}\label{sec:instrument}
